---
title: "EDA on Ecommerce Dataset"
author: "Nazir Ali Kahan"
date: "`r Sys.Date()`"
output: 
  html_document: 
    theme: lumen
    toc: yes
---

## 1. Loading all the necessary libraries

```{r}
library(tidyverse) # used for data manipulation
library(ggplot2) # used for visualizations
library(lubridate) # used for dates
```


## 2. Importing the Dataset from CSV file

```{r}
# Using read.csv command to load data
Pakistan.Largest.Ecommerce.Dataset <- read.csv("C:/Users/nazir ali khan/Downloads/Pakistan Largest Ecommerce Dataset.csv", stringsAsFactors=TRUE)
```

Loading this dataset into a new variable to keep the original copy in place:

```{r}
ecom <- Pakistan.Largest.Ecommerce.Dataset
```


## 3. Checking Data Types and the range of the values

```{r}
glimpse(ecom)
```

As we can see that R has treated order creation date, MV and working date as factors, we will have to change their data types.


## 4. Dealing with Data Types using Lubridate Package

```{r}
ecom$Working.Date <- as.character(ecom$Working.Date)
ecom$created_at <- as.character(ecom$created_at)
ecom$Customer.Since <- as.character(ecom$Customer.Since)
ecom$MV <- as.numeric(ecom$MV)
ecom$Working.Date <- mdy(ecom$Working.Date)
ecom$created_at <- mdy(ecom$created_at)
ecom$Customer.Since <- ym(ecom$Customer.Since)
# Checking new data types
class(ecom$MV)
class(ecom$Working.Date)
class(ecom$created_at)
```


## 5. Checking for any Missing Values

```{r}
sum(is.na(ecom))
```

Dataset has no missing values.


## 6. Exploratory Data Analysis

### 6.1 Number of Orders by Delivery Status

```{r}
ecom %>% select(item_id, status) %>% 
  group_by(status) %>%
  summarise(Total_Orders = n()) %>% 
  arrange(desc(Total_Orders)) -> orders_by_status
print(orders_by_status)
```


#### 6.1.1 Plot

```{r}
 orders_by_status %>% ggplot(aes(Total_Orders, status)) +
  geom_bar(stat = "identity", fill = "#0D1282") +
  labs(title = "Orders by Delivery Status", x = "# of Orders", y = "Order Status")
```


### 6.2 Total Earnings by Category

```{r}
ecom %>% select(category_name, grand_total) %>% 
  group_by(category_name) %>%
  summarise(Total_Amount_in_Lacs = sum(grand_total)/10^5) %>% 
  arrange(desc(Total_Amount_in_Lacs)) -> cat_total
# Rounding off values to 2 decimal places
cat_total$Total_Amount_in_Lacs <- round(cat_total$Total_Amount_in_Lacs,2)
print(cat_total)
```


### 6.3 Average Order Value by Category

```{r}
ecom %>% select(category_name, grand_total) %>% 
  group_by(category_name) %>%
  summarise(Avg_Order_Value = mean(grand_total)) -> cat_avg
cat_avg %>% arrange(desc(Avg_Order_Value)) -> cat_avg
cat_avg$Avg_Order_Value <- round(cat_avg$Avg_Order_Value,2)
print(cat_avg)
```


### 6.4 Number of Orders by Payment Method

```{r}
ecom %>% select(payment_method, item_id) %>% 
  group_by(payment_method) %>%
  summarise(Total_Payments = n()) %>% 
  arrange(desc(Total_Payments)) -> Payments_Frequency
print(Payments_Frequency)
```


### 6.5 Average Discount by Top 10 Payment Methods

```{r}
ecom %>% select(payment_method, discount_amount, item_id) %>% 
  group_by(payment_method) %>% 
  summarise(Avg_Discount = sum(discount_amount)/n()) %>% 
  arrange(desc(Avg_Discount)) %>% 
  head(10) -> Avg_Discount
Avg_Discount$Avg_Discount <- round(Avg_Discount$Avg_Discount,2)
print(Avg_Discount)
```


#### 6.5.1 Plot

```{r}
Avg_Discount %>% ggplot(aes(payment_method, Avg_Discount)) +
  geom_bar(stat = "identity", fill = "#0D1282" ) +
  coord_flip() + 
  labs(title = "Avg Discount by Payment Method", x = "Discount Amount", y = "Payment Method") 
```

### 6.6 Number of Customers by Month

```{r}
ecom %>% select(Customer.Since, Year) %>% 
  group_by(Customer.Since) %>% 
  summarise(Total_Customers = n()) -> customers_by_month
print(customers_by_month)
```

#### 6.6.1 Plot 
```{r}
customers_by_month %>% ggplot(aes(Customer.Since, Total_Customers)) +
  geom_point(color="#0D1282") +
  geom_line(color="#0D1282") +
  geom_smooth(method = lm, color = "#0D1282") +
  labs(title = "Customers by Year", x = "Year", y = "# of Customers") 
```


### 6.7 Number of Orders by Year

```{r}
ecom %>% select(Year, item_id) %>% 
  group_by(Year) %>% 
  summarise(Total_Order = n()) -> Orders_Per_Year
```


#### 6.7.1 Plot

```{r}
bar_chart <- Orders_Per_Year %>%ggplot(aes(Year, Total_Order)) +
  geom_bar(stat = "identity", fill = "#0D1282") +
  labs(title = "# of Order by Year", x = "Years" , y = "# of Orders")
print(bar_chart)
```


### 6.8 Sales by Fiscal Year

```{r}
ecom %>% select(FY, grand_total) %>% 
  group_by(FY) %>% 
  summarise(Total_Sales_in_Arabs = sum(grand_total/10^9)) -> sales_year
```


#### 6.8.1 Plot

```{r}
sales_year %>% ggplot(aes(FY, Total_Sales_in_Arabs)) +
  geom_bar(fill="#0D1282", stat = "identity") + 
  labs(title = "Total Sales by Fiscal Year", x= "Fiscal Year", y = "Total Sales in Arab")
```

Thank You.....
